---
title: "Reproducible Research Project 1"
author: "Joshua Chu"
date: "May 10, 2019"
output: html_document
---  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE)
```

## Loading data and libraries.
```{r}
library(ggplot2)
library(tidyr)
library(dplyr)

rawData <- read.csv("activity.csv")
df0<-rawData
df0$date<-as.Date(df0$date,format='%Y-%m-%d')
```

## Calculate daily steps and plot using a histogram.
Here I use ggplot to produce the histogram. All NAs in the dataset are not included.
```{r}
df1<-aggregate(steps~date,na.rm=TRUE,data=df0,sum)

ggplot(df1,aes(x=date)) +           
        geom_histogram(aes(y=steps),stat = 'identity',binwidth = 1,alpha=1) +
        ggtitle('Total Steps Taken Daily (NAs omitted)') +
        xlab("Date") +
        ylab("Total Steps") +
        theme(plot.title = element_text(hjust = 0.5)) +
        theme(panel.background = element_rect('white')) +
        theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
```

## Calculate and report the mean and median of the total number of steps taken per day.
A subset of the mean and median are reported.
```{r}
avg0<-aggregate(steps~date,na.rm=TRUE,data=df0,mean)
head(avg0)
med0<-aggregate(steps~date,na.rm=TRUE,data=df0,median)
head(med0)
```



## What is the average daily activity pattern?
```{r}
gpdf0<-group_by(df0,interval)

spAvg<-summarise(gpdf0,Avg=mean(steps,na.rm = TRUE))

ggplot(spAvg,aes(x=interval,y=Avg)) +
        geom_line(colour='blue',size=.5) +
        xlab("Interval") +
        ylab("Avg Steps") +
        theme(panel.background = element_rect('white')) +
        theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
```


## Inputing missing values.
1.1 Determine the total number of NAs.
```{r}
summary(df0)
```


```{r echo=FALSE}
df2<-df0
```

1.2 Use the step average to substitute for NAs. 
```{r}
df2$steps[is.na(df2$steps)]<-mean(df0$steps,na.rm = TRUE)
```

1.3 Create a new dataset and plot a histogram. 
```{r}
df3<-aggregate(steps~date,na.rm=TRUE,data=df2,sum)    

ggplot(df3,aes(x=date)) +           
        geom_histogram(aes(y=steps),stat = 'identity',binwidth = 1,alpha=1) +
        ggtitle('Total Steps Taken Daily') +
        xlab("Date") +
        ylab("Total Steps") +
        theme(plot.title = element_text(hjust = 0.5)) +
        theme(panel.background = element_rect('white')) +
        theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
```

A subset of the mean and median are reported.
```{r}
avg1<-aggregate(steps~date,na.rm=TRUE,data=df2,mean)
head(avg1)
med1<-aggregate(steps~date,na.rm=TRUE,data=df2,median)
head(med1)
```

## What is the impact of imputing missing data on the estimates of the total daily number of steps?
Here I use the base plotting system to produce the histogram. 
```{r}
par(mfrow = c(1, 2))
plot(df1, type = "h", lwd = 5,lend = "square", main = "NAs Present",xlab='Date',ylab='Total Steps')
plot(df3, type = "h", lwd = 5, lend = "square", main = "No NAs",xlab='Date',ylab='Total Steps')
```


## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels - "weekday" and "weekend".
```{r}
df2$wkday<-factor(format(df3$date, "%A"))

levels(df2$wkday)<-list(weekday = c("Monday","Tuesday","Wednesday", "Thursday","Friday"),
                        weekend =c("Saturday", "Sunday"))
```
Plot using the base plotting system.
```{r}
par(mfrow = c(2, 1))

with(df2[df2$wkday == "weekday",], plot(aggregate(steps ~ interval, FUN = mean),
                                        type = "l", main = "Weekdays"))
with(df2[df2$wkday == "weekend",], plot(aggregate(steps ~ interval, FUN = mean),
                                        type = "l", main = "Weekends"))

```

